Analysis and approximation of mixed-dimensional PDEs on 3D-1D domains coupled with Lagrange multipliers

### 2020-09-22 reply_to_reviewers.tex

\vspace{1cm}

\section*{Reviewer \#1}

In the paper under review the authors study an example of mixed dimensional PDE, i.e. a 3D problem coupled with a 1D problem. Such problems naturally appear in various applications (e.g. biomedicine and geophysics) where slender cylindrical structures are coupled with larger 3D body. The main issue in the analysis of such problem is that trace is not well defined on set of co-dimension two for weak solutions. The authors circumvent this issue by exploiting the fact that the thin structure is not 1D curve, but rather a slender 3D structure. This leads to two formulation of the problems: 3D-1D-1D and 3D-1D-2D. The former is obtained from the latter by averaging procedure. Neumann (dynamic) coupling conditions are enforced by using a Lagrange multiplier. Well-posednes is proved for the continuous and the discrete version of the problem. Two different version of the discrete case are analyzed: the case when 3D and 1D mesh matches on the interface (the conforming case) and the case when the meshes do not necessarily match (the non-conforming case). Since the saddle point approach is used, the main technical step is to prove that inf-sup condition is satisfied. In the non-conforming case the stabilization operator is constructed in order to satisfy inf-sup condition. Finally, the comparison of the numerical solution with analytical solution for benchmark problem is presented.

The result presented in the paper are interesting and to the best of my knowledge correct. I recommend the paper for publication in the SINUM.
Remarks:
\begin{enumerate}
\item Please explain what is unknown $\lambda$ in problem (1.1). I suggest including part of the Appendix from supplementary materials where $\lambda$ is introduces as a Langrange multiplier.
  \miro{
    In the revised text we newly comment that $\lambda$ can be seen as a Lagrange multiplier related to
    the coupling constraint and refer readers to the Supplementary material where the exact definition of
    $\lambda$ is given.

    Let us note that (1.1) is a ``template'' problem where we simply wish to illustrate
    that the 3$D$ and the 1$D$ problems are coupled via $\lambda\delta_{\Gamma}$ where the strength
    of the coupling, i.e. $\lambda$, is unknown. 
    }
    \item Section 2, page 2 - $\Omega$ is open? Maybe replace "in detail" with "More precisely".\\
    \fede{The standard definition of domain is used, i.e. $\Omega$ is a open and connected set. For clarity, the definition has been added to the manuscript. }
  \item P. 3. - Please provide references for $H^{1/2}_{00}$ and interpolation of weighted spaces. In addition, it would be instructive to provide an example for Riesz map S.
    \miro{
      \textit{We actually don't use $H^s$, only $H^s_{00}$. For this reason I'd like to
      simplify the text and only discuss $-\Delta u=\lambda u$ on $H^1_0(\Gamma)$.
      An advantage is that then the Riesz map is $-\Delta^{-1}$. I don't know if
      in $H^s$ we can say that $(-\Delta+I)^{-1}$ is the Riesz map.}

      \textit{
        I included refences for $H^{1/2}_{00}$ and general interpolation in
        Hilbert spaces.
      }
      The reference has been added in the revised manuscript. We have
      also simplified the exposion: definition of the space $H^s$,
      which was not used further in the analysis, was removed and only
      the space $H^s_{00}$ is introduced. Per reviewer's suggestion the
      relevant Riesz map, $-\Delta^{-1}$, is newly given in the text.
      }

    \item p.4. (l145-l50) I suggest rewriting this paragraph. First, please refer to the exact result that you are using (from [22] or [28]) and define or refer to definition of "domains having small geometric details". Second, it is not necessary to define operator norm, but enough to say that norm of $H_{\Omega}$ is uniformly bounded.\\
    \fede{We thank the reviewer for the suggestion. In the revised manuscript, the extension operator has been replaced by the harmonic extension, therefore the reference to [28] is not needed.\\
    WHAT DOES THE SECOND PART MEAN? }
    \item p4 l152. Is $H_{\Omega}$ harmonic extension? If so, please state this in previous paragraph.\\
    \fede{Yes, the definition has been added in the revised manuscript.}
  \item p8 l177-l184 - I suggest adding a figure that illustrate the assumptions on the conforming mesh.

    \miro{
      \textit{I modified the existing figure and the caption to highlight.
        Is it sufficient?}
      In the previous version we tried to illustrate the assumptions on the conforming
      mesh with Figure 5.1. This figure has newly been modified to emphasize
      correpondence of the cells of $\mathcal{T}^{\Lambda}_h$, $\mathcal{T}^{\Gamma}_h$ to
      entities of $\mathcal{T}^{\Omega}_h$.
      }

    \item p9 (4.4) Reference for these properties of $SZ_h$\\
    \fede{The reference has been added in the revised manuscript.\\

### 2020-09-22 3d1d_LM_rev.tex
almost done rev1

\usetikzlibrary{shapes, patterns, arrows}
\usetikzlibrary{calc}
\usepackage{pgfplots}
\usepgfplotslibrary{fillbetween}
\pgfplotsset{compat=1.16}

@@

In this study we consider coupled partial differential equations on domains with mixed dimensionality, in particular we address the  3D-1D case. The mathematical structure of such problems can be represented by the following formal equations:
\begin{subequations}
\label{eq:pde}
\begin{align}
\label{eq:pde1}
  -\Delta u + u + \lambda \delta_{\Lambda} &= f &\mbox{ in } \Omega,\\
\label{eq:pde2}
 d_s^2 \ud + \ud - \lambda &= g &\mbox{ on } \Lambda,\\
\label{eq:coupling}
\mathcal{T}_\Lambda u - \ud  &=  \paoloold{q} &\mbox{ on } \Lambda . 
\end{align}
\end{subequations}
Problem \eqref{eq:pde} can be described as an example of \emph{mixed dimensional PDEs}.
Here, $u$, $\ud$,  $\lambda$ are unknowns,  $\Omega$ is a bounded domain in $\R^3$, whereas $\Lambda \subset \Omega$ is a 1D structure
parameterized in terms of $s$ and $d_s$ is the derivative with respect to $s$. 
The term $\lambda\delta_{\Lambda}$ is a Dirac measure such that 
$\int_{\Omega}\lambda(x)\delta_{\Lambda}v(x)\,\mathrm{d}x=\int_{\Lambda}\lambda(t)v(t) \,\mathrm{d}t$
for a continuous function $v$ and $\mathcal{T}_\Lambda: \Omega\rightarrow\Lambda$ is a suitable restriction operator from 3D to 1D.

@@

@@

\begin{figure}
 \centering
 \includegraphics[width=0.45\textwidth]{graphics/conform_mesh.pdf}
 \includegraphics[width=0.45\textwidth]{graphics/nonconform_mesh.pdf} 
 \vspace{-10pt}
 \captionof{figure}{
   (Left)
$\Lambda$ and $\Gamma$ conforming discretization of $\Omega$
   used for \eqref{eq:problem1} and \eqref{eq:problem2}.
   (Right) Sample discretization of the benchmark geometry in the non-conforming case for \eqref{eq:problem2}. 
 }
\label{fig:mesh}
\end{figure}

>>>>

\usetikzlibrary{shapes, patterns, arrows}
\usetikzlibrary{calc}
\usepackage{pgfplots}
\usepgfplotslibrary{fillbetween}
%\pgfplotsset{compat=1.16}

@@

In this study we consider coupled partial differential equations on domains with mixed dimensionality, in particular we address the  3D-1D case. The mathematical structure of such problems can be represented by the following formal equations:
\begin{subequations}
\label{eq:pde}
\begin{align}
\label{eq:pde1}
  -\Delta u + u + \lambda \delta_{\Lambda} &= f &\mbox{ in } \Omega,\\
\label{eq:pde2}
 d_s^2 \ud + \ud - \lambda &= g &\mbox{ on } \Lambda,\\
\label{eq:coupling}
\mathcal{T}_\Lambda u - \ud  &=  \paoloold{q} &\mbox{ on } \Lambda . 
\end{align}
\end{subequations}
Problem \eqref{eq:pde} can be described as an example of \emph{mixed dimensional PDEs}.
Here, $u$, $\ud$,  $\lambda$ are unknowns,  $\Omega$ is a bounded domain in $\R^3$, whereas $\Lambda \subset \Omega$ is a 1D structure
parameterized in terms of $s$ and $d_s$ is the derivative with respect to $s$. 
The term $\lambda\delta_{\Lambda}$ is a Dirac measure such that 
$\int_{\Omega}\lambda(x)\delta_{\Lambda}v(x)\,\mathrm{d}x=\int_{\Lambda}\lambda(t)v(t) \,\mathrm{d}t$
for a continuous function $v$ and $\mathcal{T}_\Lambda: \Omega\rightarrow\Lambda$ is a suitable restriction operator from 3D to 1D.
\miro{We remark that $\lambda$ can be viewed as a Lagrange multiplier associated with the
  coupling constraint \eqref{eq:coupling}, see \S\ref{sec:derivation} for
  detailed interpretation.
}

@@

\miro{
We will frequently use inner products and norms that are weighted. The
$L_2$ and $H^1$ inner products
weighted by a scalar function 
$w$, which is strictly positive and bounded almost everywhere,
are defined as 
follows
\begin{equation*}
%(u,v)_{\Gamma, w} = 
(u,v)_{L^2(\Omega), w} = 
\int_{\Omega} w \, u \, v d\omega \ \mbox{   and   } \
%(u,v)_{\Gamma, w} = 
(u,v)_{H^1(\Omega), w} = 
\int_{\Omega}  w \, u \, v d\omega
+ \int_{\Omega} w \, \nabla u \cdot \nabla v d\omega
\end{equation*}
whereas a weighted fractional
space \miroold{$H_{00}^s(\Gamma; w)$} is defined in terms of the interpolation
of the corresponding weighted spaces (see \cite[ch. 2.1]{MR0350177}, \cite{chandler2015interpolation}). 
For the norm of such spaces, we introduce 
the Riesz map $S$ such that 
for $u, v \in H_0^1(\Gamma)$ we have 
\[
\int_{\Omega} w \nabla\left(S u\right)\cdot\nabla vd\omega  = (u,v)_{L^2(\Gamma), w}.
\]
Then $S=-\Delta^{-1}$ is a compact self-adjoint operator. 
Assuming that $\{\lambda_k\}_k$ is the set of eigenvalues, 
$\{\phi_k\}_k$ the set of eigenvectors of $S$ 
orthonormal with respect to the inner 
product $(\cdot, \cdot)_{L^2(\Gamma), w}$ and $u\in H_0^1(\Gamma)$ can be expressed 
as $u=\sum_k c_k \phi_k$ 
then
\[
\|u\|^2_{H_{00}^s(\Gamma), w} = \sum_k \lambda_k^{-s} c_k^2 .   
\]
%The space $H_{00}^s(\Gamma; w)$ is defined analogously, but with $S$ above defined in terms of the $H^1_0$ inner product.
Owing to the positivity and boundedness of $w$ the weighted spaces equal the corresponding non-weighted spaces as sets,
but their norms are different.
}

@@

\begin{figure}
 \centering
 \includegraphics[width=0.45\textwidth]{graphics/conform_mesh.pdf}
 \includegraphics[width=0.45\textwidth]{graphics/nonconform_mesh.pdf} 
 \vspace{-10pt}
 \captionof{figure}{
   (Left)
$\Lambda$ and $\Gamma$ conforming discretization of $\Omega$
   used for \eqref{eq:problem1} and \eqref{eq:problem2}.
   Each cell of $\mathcal{T}^{\Gamma}_h$ (in blue, filled marker vertices) and $\mathcal{T}^{\Lambda}_h$ (in red, filled marker vertices)
   is a facet, respectively edge, of $\mathcal{T}^{\Omega}_h$ (in black, empty square marker vertices).
   (Right) Sample discretization of the benchmark geometry in the non-conforming case for \eqref{eq:problem2}. 
 }
\label{fig:mesh}
\end{figure}

### 2020-09-22 appendix.tex
almost done rev1

\section{Derivation of the model}
This section provides a rigorous derivation of 3D-1D-1D problem \eqref{eq:problem2} and 3D-1D-2D problem \eqref{eq:problem1}. 
The steps are similar to the derivation presented in \cite{laurino_m2an}, 
however, here the coupling conditions are different,
giving rise to coupled problems featuring Lagrange multipliers. 
Precisely, the starting point is the problem arising from \emph{Dirichlet-Neumann} conditions. 
Find $\up,\uf$ s.t.:
\begin{subequations}\label{eq:dirneu}
\begin{align}
\label{eq:dirneu1}
- \Delta \up  + \up &= f  && \text{ in } \Omega_{\oplus},\\
\label{eq:dirneu2}
- \Delta \uf  + \uf &= g  && \text{ in } \Omega_\ominus,\\
\label{eq:dirneu4}
\up - \uf &= q && \text{ on }  \Gamma,\\
\label{eq:dirneu3}
\nabla\left( \up - \uf \right) \cdot \nn_{\oplus} &=0  && \text{ on } \Gamma,\\
\label{eq:dirneu5}
\up &= 0 && \text{ on } \partial \Omega\,.
\end{align}
\end{subequations}
The coupling constraints defined on $\Gamma$ involve essential or strong conditions.
Such conditions will be enforced weakly by using the method of Lagrange multipliers~\cite{MR359352}. Then, the variational formulation of problem \eqref{eq:dirneu}
is to find $\up \in H^1_{\partial\Omega}(\Omega_\oplus), \ \uf \in H^1_{\partial\Omega_\ominus\setminus\Gamma}(\Omega_\ominus), \ \lambda \in H^{-\frac 1 2}(\Gamma)$ s.t.
\begin{subequations}\label{eq:weak_dirneu}
\begin{align}
&(u_\oplus,v_\oplus)_{H^1(\Omega_\oplus)} + (u_\ominus,v_\ominus)_{H^1(\Omega_\ominus)} 
+ \langle  v_\oplus - v_\ominus, \lambda \rangle_{\Gamma} 
\\
\nonumber
&\quad = (f,v_\oplus)_{L^2(\Omega_\oplus)} + (g,v_\ominus)_{L^2(\Omega_\ominus)}
\quad \forall v_\oplus \in H^1_{\partial\Omega}(\Omega_\oplus), \ v_\ominus \in H^1_{\partial\Omega_\ominus\setminus\Gamma}(\Omega_\ominus),
\\
& \langle u_\oplus - u_\ominus, \mu \rangle_{\Gamma} = 0
\quad \forall  \mu \in H^{-\frac 1 2}(\Gamma)\,.
\end{align}
\end{subequations}
%where $\langle v, \mu \rangle_{H^{-\frac 1 2}(\Gamma)}$ denotes the duality pairing between 
%$ \mu \in H^{-\frac 1 2}(\Gamma)$ and $v \in H^{\frac 1 2}(\Gamma)$.
where $\lambda$ is the Lagrange multiplier and it is equivalent to $\nabla \uf \cdot \nn_\ominus$.

>>>>

\section{Derivation of the model}\label{sec:derivation}
This section provides a rigorous derivation of 3D-1D-1D problem \eqref{eq:problem2} and 3D-1D-2D problem \eqref{eq:problem1}. 
The steps are similar to the derivation presented in \cite{laurino_m2an}, 
however, here the coupling conditions are different,
giving rise to coupled problems featuring Lagrange multipliers. 
Precisely, the starting point is the problem arising from \emph{Dirichlet-Neumann} conditions. 
Find $\up,\uf$ s.t.:
\begin{subequations}\label{eq:dirneu}
\begin{align}
\label{eq:dirneu1}
- \Delta \up  + \up &= f  && \text{ in } \Omega_{\oplus},\\
\label{eq:dirneu2}
- \Delta \uf  + \uf &= g  && \text{ in } \Omega_\ominus,\\
\label{eq:dirneu4}
\up - \uf &= q && \text{ on }  \Gamma,\\
\label{eq:dirneu3}
\nabla\left( \up - \uf \right) \cdot \nn_{\oplus} &=0  && \text{ on } \Gamma,\\
\label{eq:dirneu5}
\up &= 0 && \text{ on } \partial \Omega\,.
\end{align}
\end{subequations}
The coupling constraints defined on $\Gamma$ involve essential or strong conditions.
Such conditions will be enforced weakly by using the method of Lagrange multipliers~\cite{MR359352}. Then, the variational formulation of problem \eqref{eq:dirneu}
is to find $\up \in H^1_{\partial\Omega}(\Omega_\oplus), \ \uf \in H^1_{\partial\Omega_\ominus\setminus\Gamma}(\Omega_\ominus), \ \lambda \in H^{-\frac 1 2}(\Gamma)$ s.t.
\begin{subequations}\label{eq:weak_dirneu}
\begin{align}
&(u_\oplus,v_\oplus)_{H^1(\Omega_\oplus)} + (u_\ominus,v_\ominus)_{H^1(\Omega_\ominus)} 
+ \langle  v_\oplus - v_\ominus, \lambda \rangle_{\Gamma} 
\\
\nonumber
&\quad = (f,v_\oplus)_{L^2(\Omega_\oplus)} + (g,v_\ominus)_{L^2(\Omega_\ominus)}
\quad \forall v_\oplus \in H^1_{\partial\Omega}(\Omega_\oplus), \ v_\ominus \in H^1_{\partial\Omega_\ominus\setminus\Gamma}(\Omega_\ominus),
\\
& \langle u_\oplus - u_\ominus, \mu \rangle_{\Gamma} = 0
\quad \forall  \mu \in H^{-\frac 1 2}(\Gamma)\,.
\end{align}
\end{subequations}
%where $\langle v, \mu \rangle_{H^{-\frac 1 2}(\Gamma)}$ denotes the duality pairing between 
%$ \mu \in H^{-\frac 1 2}(\Gamma)$ and $v \in H^{\frac 1 2}(\Gamma)$.
where $\lambda$ is the Lagrange multiplier and it is equivalent to $\nabla \uf \cdot \nn_\ominus$.
